---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# saebenchmarking

<!-- badges: start -->
[![R-CMD-check](https://github.com/fionaaudia/saebenchmarking/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/fionaaudia/saebenchmarking/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview

saebenchmarking provides model-agnostic **difference**, **ratio**, and
**optimum** benchmarking for small area estimators, together with mean
squared error (MSE) estimation, so that weighted small area aggregates
are consistent with a known external or internal target.

Point estimation methods follow You & Rao (2002), Wang, Fuller & Qu
(2008), and Rao & Molina (2015, Section 6.4.6). MSE estimation follows
Butar & Lahiri (2003) and Steorts & Ghosh (2013) for a residual
bootstrap approach, and Datta, Ghosh, Steorts & Maples (2011) for a
posterior-simulation approach suited to Hierarchical Bayes estimators.

## Installation

You can install the development version of saebenchmarking from
[GitHub](https://github.com/) with:

```r
# install.packages("pak")
pak::pak("fionaaudia/saebenchmarking")
```

## Example

### Difference benchmarking

This is a basic example showing how to benchmark small area estimates
so that their weighted aggregate matches a known external target.

```{r example}
library(saebenchmarking)

est    <- c(10.2, 8.7, 12.1)
mse    <- c(0.5, 0.4, 0.6)
weight <- c(0.3, 0.4, 0.3)

hasil <- sae_benchmarking(
  method   = "difference",
  estimate = est,
  mse      = mse,
  weight   = weight,
  target   = 10
)

hasil$benchmark$estimation
hasil$aggregation
```

### Ratio and optimum benchmarking

The same call works with `method = "ratio"` or `method = "optimum"`;
only the internal adjustment differs.

```{r ratio-optimum}
sae_benchmarking(
  method   = "ratio",
  estimate = est,
  mse      = mse,
  weight   = weight,
  target   = 10
)$benchmark$estimation

sae_benchmarking(
  method   = "optimum",
  estimate = est,
  mse      = mse,
  weight   = weight,
  target   = 10
)$benchmark$estimation
```

### Internal benchmarking (no external target)

If you don't have an external target, supply `direct` estimates
instead and the target is computed internally as their weighted sum.

```{r internal-target}
direct <- c(10.5, 8.2, 12.8)
vardir <- c(0.6, 0.5, 0.7)

hasil_internal <- sae_benchmarking(
  method   = "difference",
  estimate = est,
  mse      = mse,
  weight   = weight,
  direct   = direct,
  vardir   = vardir
)

hasil_internal$benchmark$target
```

### Estimating the MSE of benchmarked estimates

`mse_benchmarking()` takes the output of `sae_benchmarking()` and
estimates the MSE of the *benchmarked* estimates. When `direct` and
`vardir` are available, it uses a residual bootstrap by default:

```{r mse-bootstrap}
mse_benchmarking(hasil_internal, B = 200, seed = 1)
```

For Hierarchical Bayes estimators, supply a matrix of posterior draws
instead, and the posterior-simulation approach is used:

```{r mse-posterior}
set.seed(1)
posterior_draws <- matrix(
  rnorm(300 * 3, rep(est, each = 300), sqrt(mse)),
  ncol = 3
)

mse_benchmarking(hasil, posterior_samples = posterior_draws)
```

## Learn more

See `vignette("pengantar-benchmarking")` for a more detailed
walkthrough, including guidance on choosing between the three
benchmarking methods and between the bootstrap and posterior MSE
approaches.
