---
title: "Introduction to saebenchmarking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to saebenchmarking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(saebenchmarking)
```

## Why benchmarking?

Model-based small area estimates, such as the EBLUP under the Fay-Herriot
model, usually do not add up to the reliable direct estimate of the larger
area. Benchmarking adjusts them so that
$\sum_i w_i \hat\theta_i^{B} = \sum_i w_i \hat\theta_i^{DIR}$.

## Data

`data_eblup` contains simulated data for 50 areas with EBLUP estimates;
`data_hb` contains the same areas with hierarchical Bayes estimates.

```{r}
head(data_eblup[, c("area", "weight", "direct", "vardir", "eblup", "mse")])
```

## Point benchmarking

```{r}
methods <- c("difference", "ratio", "optimum")
bench <- lapply(methods, function(m) {
  sae_benchmarking(method = m, direct = "direct", weight = "weight",
                   estimate = "eblup", mse = "mse", vardir = "vardir",
                   phi_source = "mse", data = data_eblup)
})
names(bench) <- methods

sapply(bench, function(x) x$aggregation[, c("Estimate", "Bench", "Target")])
```

The difference method shifts every area by the same amount, the ratio method
scales every area by the same factor, and the optimum method distributes the
discrepancy according to the MSE (or sampling variance) of each area.

## MSE of benchmarked EBLUPs

```{r}
Z   <- cbind(1, data_eblup$z)
s2v <- attr(data_eblup, "sigma2_v")

mse_diff <- mse_benchmarking(bench$difference, estimator = "eblup",
                             z = Z, sigma2_v = s2v, fitting_method = "REML")

mse_opt <- mse_benchmarking(bench$optimum, estimator = "eblup",
                            z = Z, sigma2_v = s2v, B = 100, seed = 2026)

comparison <- data.frame(
  EBLUP      = data_eblup$mse,
  Difference = mse_diff$mse,
  Optimum    = mse_opt$mse
)
head(round(comparison, 4))
```

A small number of bootstrap replicates is used here to keep the vignette
fast; use a larger `B` (for example the default of 1000) in practice.

## Posterior MSE of benchmarked HB estimates

```{r}
bm_hb <- sae_benchmarking("optimum", direct = "direct", weight = "weight",
                          estimate = "theta_hb", vardir = "vardir",
                          phi_source = "vardir", data = data_hb)

mse_hb <- mse_benchmarking(bm_hb, estimator = "hb",
                           theta_hb = data_hb$theta_hb, V_hb = data_hb$var_hb)
head(data.frame(V_hb = data_hb$var_hb, PMSE = mse_hb$mse))
```

## References

Datta, G. S., Ghosh, M., Steorts, R. and Maples, J. (2011). Bayesian
benchmarking with applications to small area estimation. *TEST*, 20(3),
574-588.

Rao, J. N. K. and Molina, I. (2015). *Small Area Estimation*, 2nd edition.
Wiley.

Steorts, R. C. and Ghosh, M. (2013). On estimation of mean squared errors of
benchmarked empirical Bayes estimators. *Statistica Sinica*, 23(2), 749-767.

Sugasawa, S., Tamae, H. and Kubokawa, T. (2017). Bayesian estimators for
small area models shrinking both means and variances. *Scandinavian Journal
of Statistics*, 44(1), 150-167.

Wang, J., Fuller, W. A. and Qu, Y. (2008). Small area estimation under a
restriction. *Survey Methodology*, 34(1), 29-36.
